Parsing Postfix log files

### 2008-05-20 doc/logparser-other-parsers-review.tex
:%s/logs/log files/gc

    \item \parsername{} does not provide any reporting based on the data it
        gathers.  This is by design: the parser is responsible for parsing
        logs and extracting data --- further processing of that data is a
        separate concern, deferred to another program.  Using an \SQL{}
        database simplifies the process of generating such reports
        (discussed in \sectionref{database as API}); some
        simple examples are given in \sectionref{motivation}.

@@

    \item A minor difference is that most parsers don't handle compressed
        files; both \parsername{} and Splunk handle them transparently,
        without user intervention; Sawmill and Lire can be configured to
        support compressed files, but Sawmill exhibits a dramatic slow down
        when thus configured.  Although this is a minor difference, support
        for reading compressed log files is quite helpful, as it
        dramatically reduces the disk space required to store old logs.

@@

Pflogsumm produces a report designed for troubleshooting, rather than for
in-depth analysis.  It does not support saving any data, nor does it
extract more data than is required for producing the report.  Both the
parsing and reporting are difficult to extend as it is a specialised tool,
unlike the easily extensible design of \parsername{}.  It does not attempt
to correlate log lines by queueid or \pid{}, nor does it need to deal with
the complications encountered during this project.  Pflogsumm produces a
useful report, and successfully dealt with the \numberOFlogFILES{} logs
files tested with,\footnote{The results it reported were not verified, but
it did not report any errors, and has a very good reputation amongst
Postfix users.} but is not a suitable base for this project.

@@

Splunk aims to index all an organisation's logs, providing a centralised
view capable of searching and correlating diverse log sources.  The web
interface supports complicated searches, providing statistics and graphs in
real time, a facility not provided by \parsername{} (report generation has
been deferred to a subsequent program).  Saved searches\footnote{The author
was unable to save searches, though that may have been due to limitations
in the free version.} can be run periodically and the results emailed to a
recipient or sent to a shell script, which presumably can publish the
results as required (though possibly without the graphs and detailed
statistics); \parsername{} provides the database and leaves it to the user
to utilise it in any way, whenever they want, with no restrictions on usage
of the data.  The interface is optimised for interactive rather than
automated queries and it does not appear to be possible to write
independent tools to utilise the Splunk database; there are additional
reports available on \url{http://www.splunkbase.com/}, but the author was
unable to find any: every category was empty, even those which the
interface claimed had reports available.  Many types of reports are bundled
with the software, though most are variations of a bar or pie chart, with
the exception of bubble and heatmap graphs.  It is very easy to drill down
through the graphs to extract a portion of the data (e.g.\ select the hour
with the largest number of events, then select a particular host, and
finally a specific address), though it is not possible to search on partial
words.

@@

In the specific case of parsing Postfix logs, Splunk extracts some standard
fields: to and from addresses, HELO hostname, date, host the logs were
collected from, and protocol.  It parses the standard syslog fields at the
beginning of the line, and extracts any \texttt{key=<value>} pairs
occurring after the standard syslog prologue; these pairs are the fields
listed above.  Searches can be based on the extracted fields, and all the
text in the line is also available for searching.  \parsername{} extracts
noticeably more data, though it does not make the full text of the line
available, and the full power of \SQL{} is available when searching,
allowing the user to search on arbitrarily complicated conditions.

@@

AWStats will produce simple graphs for many different services, but
supporting many different services without special purpose code restricts
it to supporting the \LCD{}.  The data it will extract from an \MTA{} log
file is limited in comparison to \parsername{}: \newline \tab{} time2,
email, email\_r, host, host\_r, method, url, code and bytesd.\newline There
does not seem to be an explanation of any of those fields in the
documentation (\parsername{} provides copious documentation).  AWStats has
no special purpose code to deal with the intricacies of Postfix logs, in
fact it operates by coercing Postfix logs into Apache\footnote{The Apache
web server is the most popular HTTP server in use for the past 10 years;
more information is available at \url{http://httpd.apache.org/}.} format
log files, for analysis by AWStats' HTTP log file parser.  The converting
parser only deals with a small portion of the log lines generated by
Postfix, silently skipping those it cannot deal with, and does not
distinguish between different types of rejection; there is no easy way to
extend it to handle new log lines.  Although though it does correlate log
lines by queueid, it does not deal with any of the other complications
described in this document.  AWStats supports saving data but the format of
the saved data is not documented (as far as the author could tell).  It
also supports reading compressed log files, but that functionality was not
tested.

@@

This utility tracks the number of events which occurred over a particular
time and warns if the frequency of events passes a certain threshold.  It's
designed to provide real time alerts when dictionary attacks, mail loops or
similar problems occur. It doesn't attempt to extract or save data,
correlate log lines, or any of the more advanced tasks described in this
document, because it is not designed to do so.  The user must construct the
\regexes{} to match significant input lines before it will parse any logs.
This program was not reviewed in further detail because its aims are so far
removed from the aims of this project.

@@

Lire is a general purpose log parser supporting many different types of log
file.  Its Postfix parser extracts the following data from Postfix logs:
\textit{The email servers' reports will show you the number of deliveries
and the volume of email delivered by day, the domains from which you
receive or send the most emails, the relays most used, etc.\/}; notably
rejections are not mentioned or dealt with, and unlike \parsername{} there
is no facility to extend the parser.  It supports multiple output formats
for generated reports (text, \HTML{}, \PDF{} and Excel 95) but the reports
do not appear to be customisable; \parsername{} does not produce any
reports.  Lire supports saving extracted data for later report generation,
but accessing this data from another application is undocumented; given the
source code it should be possible, with sufficient time and effort, to
access the data from an external program.

@@

Logrep extracts far less data from Postfix logs than \parsername{},
completely ignores rejections, is effectively undocumented, doesn't deal
with the more complicated aspects of Postfix logs, and at the time of
writing doesn't work properly.

@@

While there are other programs available which perform basic Postfix log
parsing (some to a greater level of detail than others), few attempt to
correlate log lines by queueid (none correlate by \pid{}) to produce an
overall record of the journey each mail traverses through Postfix.  None of
the reviewed parsers collect the breadth of information gathered by
\parsername{}, nor make it as easy to extend the parser to handle new log
lines.  Most other parsers immediately generate a report and discard the
data extracted from the logs; those which don't discard the data retain it
in a format inaccessible to other tools.  Nearly all of the parsers
reviewed can produce a report of greater or lesser detail and complexity, a
facility not offered by \parsername{}; reporting is deferred to a
subsequent application.  The quality of the documentation offered by the
subset of parsers which provide some varies from unusable to good; none of
the parsers provide any documentation on the format of their data stores
(if any).  Less than half of the parsers were capable of parsing the
\numberOFlogFILES{} test log files, and improving or extending the parsers
would have been quite a difficult task for any of the parsers.  Table
\refwithpage{Summary of parsers' features} provides a summary of the
parsers' features.

The overriding difference between \parsername{} and the other parsers
reviewed herein is that none of them aim to achieve the high level of
understanding of Postfix logs achieved by \parsername{}.

>>>>

    \item \parsername{} does not provide any reporting based on the data it
        gathers.  This is by design: the parser is responsible for parsing
        log files and extracting data --- further processing of that data
        is a separate concern, deferred to another program.  Using an
        \SQL{} database simplifies the process of generating such reports
        (discussed in \sectionref{database as API}); some simple examples
        are given in \sectionref{motivation}.

@@

    \item A minor difference is that most parsers don't handle compressed
        files; both \parsername{} and Splunk handle them transparently,
        without user intervention; Sawmill and Lire can be configured to
        support compressed files, but Sawmill exhibits a dramatic slow down
        when thus configured.  Although this is a minor difference, support
        for reading compressed log files is quite helpful, as it
        dramatically reduces the disk space required to store old log
        files.

@@

Pflogsumm produces a report designed for troubleshooting, rather than for
in-depth analysis.  It does not support saving any data, nor does it
extract more data than is required for producing the report.  Both the
parsing and reporting are difficult to extend as it is a specialised tool,
unlike the easily extensible design of \parsername{}.  It does not attempt
to correlate log lines by queueid or \pid{}, nor does it need to deal with
the complications encountered during this project.  Pflogsumm produces a
useful report, and successfully dealt with the \numberOFlogFILES{} log
files files tested with,\footnote{The results it reported were not
verified, but it did not report any errors, and has a very good reputation
amongst Postfix users.} but is not a suitable base for this project.

@@

Splunk aims to index all an organisation's log files, providing a
centralised view capable of searching and correlating diverse log sources.
The web interface supports complicated searches, providing statistics and
graphs in real time, a facility not provided by \parsername{} (report
generation has been deferred to a subsequent program).  Saved
searches\footnote{The author was unable to save searches, though that may
have been due to limitations in the free version.} can be run periodically
and the results emailed to a recipient or sent to a shell script, which
presumably can publish the results as required (though possibly without the
graphs and detailed statistics); \parsername{} provides the database and
leaves it to the user to utilise it in any way, whenever they want, with no
restrictions on usage of the data.  The interface is optimised for
interactive rather than automated queries and it does not appear to be
possible to write independent tools to utilise the Splunk database; there
are additional reports available on \url{http://www.splunkbase.com/}, but
the author was unable to find any: every category was empty, even those
which the interface claimed had reports available.  Many types of reports
are bundled with the software, though most are variations of a bar or pie
chart, with the exception of bubble and heatmap graphs.  It is very easy to
drill down through the graphs to extract a portion of the data (e.g.\
select the hour with the largest number of events, then select a particular
host, and finally a specific address), though it is not possible to search
on partial words.

@@

In the specific case of parsing Postfix log files, Splunk extracts some
standard fields: to and from addresses, HELO hostname, date, host the log
files were collected from, and protocol.  It parses the standard syslog
fields at the beginning of the line, and extracts any \texttt{key=<value>}
pairs occurring after the standard syslog prologue; these pairs are the
fields listed above.  Searches can be based on the extracted fields, and
all the text in the line is also available for searching.  \parsername{}
extracts noticeably more data, though it does not make the full text of the
line available, and the full power of \SQL{} is available when searching,
allowing the user to search on arbitrarily complicated conditions.

@@

AWStats will produce simple graphs for many different services, but
supporting many different services without special purpose code restricts
it to supporting the \LCD{}.  The data it will extract from an \MTA{} log
file is limited in comparison to \parsername{}: \newline \tab{} time2,
email, email\_r, host, host\_r, method, url, code and bytesd.\newline There
does not seem to be an explanation of any of those fields in the
documentation (\parsername{} provides copious documentation).  AWStats has
no special purpose code to deal with the intricacies of Postfix log files,
in fact it operates by coercing Postfix log files into Apache\footnote{The
Apache web server is the most popular HTTP server in use for the past 10
years; more information is available at \url{http://httpd.apache.org/}.}
format log files, for analysis by AWStats' HTTP log file parser.  The
converting parser only deals with a small portion of the log lines
generated by Postfix, silently skipping those it cannot deal with, and does
not distinguish between different types of rejection; there is no easy way
to extend it to handle new log lines.  Although though it does correlate
log lines by queueid, it does not deal with any of the other complications
described in this document.  AWStats supports saving data but the format of
the saved data is not documented (as far as the author could tell).  It
also supports reading compressed log files, but that functionality was not
tested.

@@

This utility tracks the number of events which occurred over a particular
time and warns if the frequency of events passes a certain threshold.  It's
designed to provide real time alerts when dictionary attacks, mail loops or
similar problems occur. It doesn't attempt to extract or save data,
correlate log lines, or any of the more advanced tasks described in this
document, because it is not designed to do so.  The user must construct the
\regexes{} to match significant input lines before it will parse any log
files.  This program was not reviewed in further detail because its aims
are so far removed from the aims of this project.

@@

Lire is a general purpose log parser supporting many different types of log
file.  Its Postfix parser extracts the following data from Postfix log
files: \textit{The email servers' reports will show you the number of
deliveries and the volume of email delivered by day, the domains from which
you receive or send the most emails, the relays most used, etc.\/}; notably
rejections are not mentioned or dealt with, and unlike \parsername{} there
is no facility to extend the parser.  It supports multiple output formats
for generated reports (text, \HTML{}, \PDF{} and Excel 95) but the reports
do not appear to be customisable; \parsername{} does not produce any
reports.  Lire supports saving extracted data for later report generation,
but accessing this data from another application is undocumented; given the
source code it should be possible, with sufficient time and effort, to
access the data from an external program.

@@

Logrep extracts far less data from Postfix log files than \parsername{},
completely ignores rejections, is effectively undocumented, doesn't deal
with the more complicated aspects of Postfix log files, and at the time of
writing doesn't work properly.

@@

While there are other programs available which perform basic Postfix log
parsing (some to a greater level of detail than others), few attempt to
correlate log lines by queueid (none correlate by \pid{}) to produce an
overall record of the journey each mail traverses through Postfix.  None of
the reviewed parsers collect the breadth of information gathered by
\parsername{}, nor make it as easy to extend the parser to handle new log
lines.  Most other parsers immediately generate a report and discard the
data extracted from the log files; those which don't discard the data
retain it in a format inaccessible to other tools.  Nearly all of the
parsers reviewed can produce a report of greater or lesser detail and
complexity, a facility not offered by \parsername{}; reporting is deferred
to a subsequent application.  The quality of the documentation offered by
the subset of parsers which provide some varies from unusable to good; none
of the parsers provide any documentation on the format of their data stores
(if any).  Less than half of the parsers were capable of parsing the
\numberOFlogFILES{} test log files, and improving or extending the parsers
would have been quite a difficult task for any of the parsers.  Table
\refwithpage{Summary of parsers' features} provides a summary of the
parsers' features.

The overriding difference between \parsername{} and the other parsers
reviewed herein is that none of them aim to achieve the high level of
understanding of Postfix log files achieved by \parsername{}.

### 2008-05-20 doc/logparser-parsing-algorithm.tex
:%s/logs/log files/gc

Where the rules are quite simple and each rule is completely independent of
its fellows, the algorithm is significantly more complicated and highly
internally interdependent.  The algorithm deals with all the complications
of parsing, the eccentricities and oddities of Postfix logs, presenting the
resulting data in a normalised, easy to use representation.  The
algorithm's task is to follow the journey each mail takes through Postfix,
combining the data extracted by rules into a coherent whole, saving it in a
useful and consistent form, and performing housekeeping duties.

@@

The most difficult complication initially encountered is that locally
addressed mails are not always delivered directly to a mailbox: sometimes
they are addressed to and accepted for a local address but need to be
delivered to one or more remote addresses due to aliases.  When this occurs
a child mail will be injected into the Postfix queue, but without the
explicit logging \daemon{smtpd} or \daemon{postdrop} injected mails have.
Thus the source is not immediately discernible from the log line in which
the mail first appears; from a strictly chronological reading of the logs
it usually appears as if the child mail has appeared from thin air.
Subsequently the parent mail will log the creation of the child mail:

@@

The crux of the problem is that re-injected mails appear in the logs
without explicit logging indicating their source.  There are two implicit
indications:

@@

Adding new actions is not as simple as adding new rules, though care has
been taken in the parser's design and implementation to make adding new
actions as painless as possible.  The implementor writes a subroutine which
accepts the standard arguments given to actions, and registers it as an
action\footnote{The new action must be registered before the rules are
loaded, as it is an error for a rule to reference an unregistered action;
this helps catch mistakes made when adding new rules.} by calling the
framework subroutine add\_actions() with the name of the new action
subroutine as a parameter.  No other work is required on the implementor's
part to integrate the action into the parser; all of their attention and
effort can be focused on the correctness of their action.  The only
negative aspect is that the process involves editing the parser source
code, which makes upgrading to a later version of the parser more
difficult, though by no means impossible.  If the author of the new action
wishes, they can take advantage of the parser's object oriented
implementation~\cite{Wikipedia-object-orientation} by subclassing it and
implementing their changes in the derived class, allowing future upgrading
of the parser with greatly reduced chance of conflicts.\footnote{The real
difference between the two approaches is where the new code is placed.  The
simpler option is to change the parser code directly, but those changes
will then have to be made to subsequent versions of the parser, and as the
scope of the changes increases so does the chance of conflict, or mistakes
when copying the action.  The more time consuming option is to write a
subclass containing the new actions and change the program which invokes
the parser so that it uses the subclass rather than the parser; the changes
required to the program invoking the parser are minor and much less likely
to lead to conflicts when upgrading to future versions of the parser.  An
alternative is to submit new actions to the author of the parser for
inclusion in future versions, resulting in two benefits: the new actions do
not need to be maintained separately, and other users of the parser will
avail of the new functionality.} The action may need to extend the list of
valid combinations described in \sectionref{out of order log lines} if the
addition creates a different set of acceptable programs, but this is
unlikely to occur, as it would require parsing log lines from Postfix
components the parser currently ignores.\footnote{The mail server used for
development does not utilise either the \daemon{lmtp} or \daemon{virtual}
delivery agents, so this parser does not have rules to handle log lines
from those components.  Adding new rules to parse those component's log
lines is a simple process, though if their behaviour differs significantly
from the \daemon{smtp} or \daemon{local} delivery agents new actions may be
required.  The mail server in question is a production mail server handling
mail for a university department; the benefit is that the logs used exhibit
the idiosyncrasies and peculiarities a mail server in the wild must deal
with, but the downside is that significantly altering the configuration
just to log messages from a different Postfix component is not an option.}

@@

Occasionally a \daemon{smtpd} will just stop logging, without an
immediately obvious reason.  After poring over logs for some time there are
several reasons for this infrequent occurrence:

>>>>

Where the rules are quite simple and each rule is completely independent of
its fellows, the algorithm is significantly more complicated and highly
internally interdependent.  The algorithm deals with all the complications
of parsing, the eccentricities and oddities of Postfix log files,
presenting the resulting data in a normalised, easy to use representation.
The algorithm's task is to follow the journey each mail takes through
Postfix, combining the data extracted by rules into a coherent whole,
saving it in a useful and consistent form, and performing housekeeping
duties.

@@

The most difficult complication initially encountered is that locally
addressed mails are not always delivered directly to a mailbox: sometimes
they are addressed to and accepted for a local address but need to be
delivered to one or more remote addresses due to aliases.  When this occurs
a child mail will be injected into the Postfix queue, but without the
explicit logging \daemon{smtpd} or \daemon{postdrop} injected mails have.
Thus the source is not immediately discernible from the log line in which
the mail first appears; from a strictly chronological reading of the log
files it usually appears as if the child mail has appeared from thin air.
Subsequently the parent mail will log the creation of the child mail:

@@

The crux of the problem is that re-injected mails appear in the log files
without explicit logging indicating their source.  There are two implicit
indications:

@@

Adding new actions is not as simple as adding new rules, though care has
been taken in the parser's design and implementation to make adding new
actions as painless as possible.  The implementor writes a subroutine which
accepts the standard arguments given to actions, and registers it as an
action\footnote{The new action must be registered before the rules are
loaded, as it is an error for a rule to reference an unregistered action;
this helps catch mistakes made when adding new rules.} by calling the
framework subroutine add\_actions() with the name of the new action
subroutine as a parameter.  No other work is required on the implementor's
part to integrate the action into the parser; all of their attention and
effort can be focused on the correctness of their action.  The only
negative aspect is that the process involves editing the parser source
code, which makes upgrading to a later version of the parser more
difficult, though by no means impossible.  If the author of the new action
wishes, they can take advantage of the parser's object oriented
implementation~\cite{Wikipedia-object-orientation} by subclassing it and
implementing their changes in the derived class, allowing future upgrading
of the parser with greatly reduced chance of conflicts.\footnote{The real
difference between the two approaches is where the new code is placed.  The
simpler option is to change the parser code directly, but those changes
will then have to be made to subsequent versions of the parser, and as the
scope of the changes increases so does the chance of conflict, or mistakes
when copying the action.  The more time consuming option is to write a
subclass containing the new actions and change the program which invokes
the parser so that it uses the subclass rather than the parser; the changes
required to the program invoking the parser are minor and much less likely
to lead to conflicts when upgrading to future versions of the parser.  An
alternative is to submit new actions to the author of the parser for
inclusion in future versions, resulting in two benefits: the new actions do
not need to be maintained separately, and other users of the parser will
avail of the new functionality.} The action may need to extend the list of
valid combinations described in \sectionref{out of order log lines} if the
addition creates a different set of acceptable programs, but this is
unlikely to occur, as it would require parsing log lines from Postfix
components the parser currently ignores.\footnote{The mail server used for
development does not utilise either the \daemon{lmtp} or \daemon{virtual}
delivery agents, so this parser does not have rules to handle log lines
from those components.  Adding new rules to parse those component's log
lines is a simple process, though if their behaviour differs significantly
from the \daemon{smtp} or \daemon{local} delivery agents new actions may be
required.  The mail server in question is a production mail server handling
mail for a university department; the benefit is that the log files used
exhibit the idiosyncrasies and peculiarities a mail server in the wild must
deal with, but the downside is that significantly altering the
configuration just to log messages from a different Postfix component is
not an option.}

@@

Occasionally a \daemon{smtpd} will just stop logging, without an
immediately obvious reason.  After poring over log files for some time
there are several reasons for this infrequent occurrence:

### 2008-05-20 doc/logparser-parsing-rules.tex
:%s/logs/log files/gc

Graphs~\refwithpage{percentage increase of shuffled over normal}
and~\refwithpage{percentage increase of reversed over normal} show the
percentage increase of execution times, with table~\refwithpage{Execution
time increase for different rule orderings} showing the mean increases for
different groupings of logs.  Overall this provides a modest but worthwhile
performance increase of approximately 9\%, for a small investment in time
and programming.

@@

The logs produced by Postfix differ from installation to installation,
because administrators have the freedom to choose the subset of available
restrictions which suits their needs, including using different \RBL{}
services, policy servers, or custom rejection messages.  To facilitate easy
parsing of new log lines, the parser's design separates parsing rules from
parsing actions: adding new actions is difficult, but adding new rules to
parse new rejection messages is trivial (and also occurs much more
frequently).  The implementation provides a program to ease the process of
creating new rules from unparsed log lines, based on the algorithm
developed by Risto Vaarandi~\cite{risto-vaarandi} for his
\SLCT{}~\cite{slct-paper}.  The differences between the two algorithms will
be outlined as part of the general explanation below.

>>>>

Graphs~\refwithpage{percentage increase of shuffled over normal}
and~\refwithpage{percentage increase of reversed over normal} show the
percentage increase of execution times, with table~\refwithpage{Execution
time increase for different rule orderings} showing the mean increases for
different groupings of log files.  Overall this provides a modest but
worthwhile performance increase of approximately 9\%, for a small
investment in time and programming.

@@

The log files produced by Postfix differ from installation to installation,
because administrators have the freedom to choose the subset of available
restrictions which suits their needs, including using different \RBL{}
services, policy servers, or custom rejection messages.  To facilitate easy
parsing of new log lines, the parser's design separates parsing rules from
parsing actions: adding new actions is difficult, but adding new rules to
parse new rejection messages is trivial (and also occurs much more
frequently).  The implementation provides a program to ease the process of
creating new rules from unparsed log lines, based on the algorithm
developed by Risto Vaarandi~\cite{risto-vaarandi} for his
\SLCT{}~\cite{slct-paper}.  The differences between the two algorithms will
be outlined as part of the general explanation below.

### 2008-05-20 doc/logparser.tex
Tested Splunk.
Tested Isoqlog and Anteater. Didn't test "Log analyser --- throughput monitor" - it would be pointless.
Tested lire. Logrep can't be tested, it doesn't work properly.
Improved other parsers introduction. Sawmill tested. YAALA tested.
Added summary table to other parsers' conclusion. Some more details about AWStats. Started testing LMA.
Added cluster_groups to the rules table description and example rule. Improved Splunk review. Improved summary table in review of other parsers conclusion. Minor improvements in parser reviews. Improved review of other parsers conclusion.
Moved \a

    Parsing Postfix logs is much more difficult than it first appears, but
    it \textit{is\/} possible to achieve a high degree of accuracy in
    understanding the logs, and thus accuracy in reconstructing the actions
    taken by Postfix to generate the logs.  This paper describes the
    creation of a parser, documenting the parsing algorithm, explaining the
    difficulties encountered and the solutions developed, with reference to
    an implementation which stores data gleaned from the logs in an SQL
    database.  The gathered data can then be used to optimise current
    anti-spam measures, provide a baseline to test new anti-spam measures
    against, or to produce statistics showing how effective those measures
    are.

@@

XXX TEST OTHER PARSERS ON THE 93 TEST LOG FILES\@:

\begin{enumerate}

    \item Sawmill Universal Log File Analysis and Reporting
    \item Splunk
    \item Isoqlog
    \item Log analyser --- throughput monitor
    \item Anteater
    \item Yet Another Advanced Logfile Analyser
    \item Logparser/Lire
    \item Logrep
    \item Log Mail Analyser

\end{enumerate}

@@

\newpage
\input{logparser-introduction.tex}
\input{logparser-background.tex}
\input{logparser-database-schema.tex}
\input{logparser-parsing-rules.tex}
\input{logparser-parsing-algorithm.tex}
\input{logparser-coverage.tex}
\input{logparser-limitations.tex}
\input{logparser-conclusion.tex}
\input{logparser-other-parsers-review.tex}

>>>>

    Parsing Postfix log files is much more difficult than it first appears,
    but it \textit{is\/} possible to achieve a high degree of accuracy in
    understanding the log files, and thus accuracy in reconstructing the
    actions taken by Postfix to generate the log files.  This paper
    describes the creation of a parser, documenting the parsing algorithm,
    explaining the difficulties encountered and the solutions developed,
    with reference to an implementation which stores data gleaned from the
    log files in an SQL database.  The gathered data can then be used to
    optimise current anti-spam measures, provide a baseline to test new
    anti-spam measures against, or to produce statistics showing how
    effective those measures are.

@@

XXX WEKA FOR DECISION TREE DISPLAY\@.

@@

\newpage
\input{logparser-introduction.tex}
\input{logparser-background.tex}
\input{logparser-database-schema.tex}
\input{logparser-parsing-rules.tex}
\input{logparser-parsing-algorithm.tex}
\input{logparser-coverage.tex}
\input{logparser-limitations.tex}
\input{logparser-conclusion.tex}

\appendix

\input{logparser-other-parsers-review.tex}
